A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-04-23 doc/paper.tex
User study. Fixes #71, fixes #72, fixes #73, fixes #74, fixes #75, fixes #76, fixes #77.

	\subsection{Bad Smells} Fowler's description of code smells
	\cite{fowler2009refactoring} underscored an important point of code quality:
	between perfect code and bug-crippled spaghetti,  there is a spectrum of code
	designs which, by themselves, are not faulty but nevertheless suggest vulnerabilities in code design. 	
	Since then, researchers have applied these concepts to spreadsheet programming, as 
	discussed in section \ref{related-work}. While some of the smells lie beyond the scope
	of the tool, such as those characterizing inter-worksheet connections, others
	have structures which create distinctive features within the tree, leading to
	easy detection through our visualization: \par

@@

		\item \textit{Long Method}: Also called "Multiple Operations"
		in \cite{hermans2012detecting}, this refers to situations where numerous
		functions and operators are used within a single formula. As a result, the formula
		tends to be prohibitively complex. In the visualization, since each nested and combined function
		creates yet another level of depth within the tree, these constructions can
		result in long, horizontal chains of nodes, as seen in figure
		\ref{"fig:longsum"}.

		\item \textit{Long Parameter List}: [tall, vertical towers
		of blue nodes]

		\item \textit{Conditional Complexity}: [IF nodes within IF
		nodes] \end{itemize}

@@

	Some examples are meaningful not for the space they take up but for how they
	clash with the expectations around a function. For example, Excel offers a
	number of functions that accept any number arguments given to them, up to a
	resource-defined limit. SUM and CONCATENATE are two such functions: adding more
	arguments simply adds or appends that element in the same way as every previous
	argument. Strange cases were found, then, in the cases where these functions
	were used with only one argument. \par

@@

	\section{User Study} Midway through production, we conducted a brief,
	four-participant user study in order to gauge how well we were capturing the
	initial design goals. The study was, in general, a flexible and exploratory
	affair: rather than giving the users any specific tasks or pointed questions,
	we asked instead that they choose for themselves which trees to explore and
	report anything that they deemed interesting. Specifically, given trees rooted
	in the Enron dataset and at least 20 to thirty minutes, the participants
	situated themselves within the persona of a consultant asked to evaluate a
	company's spreadsheet practice, which could be arranged around questions like
	which functions on which the employees most depended, which sheets presented
	the most anomalous or dangerous designs, and so on. Interpretation of
	"spreadsheet practice" can certainly vary from participant to participant, but
	since we wanted a range of views and backgrounds despite the persona, we
	readily accepted this interpretive wiggle room. \par

@@

	\begin{wrapfigure}{l}{.12\textwidth} \centering \includegraphics{SUM}
		\label{fig:sum} \end{wrapfigure}

@@

	\begin{figure} \centering \includegraphics[scale=.5]{longIF} \label{fig:longif}
		\caption{The many possibilities of IF, argument 3, demonstrate the problem of
			having too many options.} \end{figure}

@@

	\includegraphics[width = .46\textwidth]{comparison}

>>>>

	\subsection{API Improvement}

	\subsection{Bad Smells} Fowler's description of code smells
	\cite{fowler2009refactoring} underscored an important point of code quality:
	between elegant, efficient code and bug-crippled spaghetti, there is a spectrum of code
	designs which, by themselves, are not faulty but nevertheless suggest vulnerabilities in code design. 	
	Since then, researchers have applied these concepts to spreadsheet programming, as 
	discussed in section \ref{related-work}. While some of the smells lie beyond the scope
	of the tool, such as those characterizing inter-worksheet connections, others
	have structures which create distinctive features within the tree, leading to
	easy detection through our visualization: \par

@@

		\item \textit{Multiple Operations}: A single formula comprises numerous
		functions and operators. \cite{hermans2012detecting}. As a result, the formula
		tends to be prohibitively complex. In the visualization, since each nested and
		combined function creates yet another level of depth within the tree, these
		constructions can result in long, horizontal chains of nodes, as seen in
		figure \ref{"fig:longsum"}.

		\item \textit{Long Parameter List}: A function uses numerous arguments, making
		it harder to understand \cite{asavametha2012detecting}. This is not a problem
		for many functions which have a limited number of arguments that it accepts. However,
		for functions that accept arguments indefinitely, such as SUM and CONCATENATE, it creates
		tall towers of argument nodes from a single function node.

		\item \textit{Conditional Complexity}: A logical function, particularly IF, nests
		several other logical functions inside of it, reducing code readability \cite{hermans2012detecting}. 
		These can be traced manually in the trees for the logical functions. IF, for example,
		will have several more IF function nodes within its second or third arguments, which
		in turn might have more of the same nested within them.

@@

	Some code issues are meaningful not for the space they occupy but for how they
	clash with a function's expectations. For example, Excel offers a
	number of functions that accept any number arguments given to them, up to a
	resource-defined limit. SUM, for example, simply adds every argument given to it.
	This also means that it can accept only one argument, and when the user does so,
	they pass in a range of cells over 90\% of the time. However, using the tool,
	we isolated an instance where the programmer used a single-argument SUM where the 
	argument was a series of 32 contiguous cells added with the + operation -- 
	``SUM(C6+C7+C8+C9+C10+C11+C12+C13+...+C37)" -- representing a technically
	valid but unusual and problematic design.\par

@@

	\section{User Study} During development, we conducted a brief, four-participant
	user study in order to detect conflicts between our design and the goals in
	section \ref{goals}. We designed the study to be exploratory; given trees based
	in the Enron dataset and at least twenty minutes, we asked the participants to
	adopt the persona of a consultant evaluating a company's spreadsheet
	practice, allowing for questions like which functions on which the employees
	most depended or which sheets or employees presented the most anomalous or
	dangerous designs. Though we instructed them on how to use the tool and
	suggested that they visit one of the most populous trees in the set, we put
	aside any specific tasks or questions and let them choose their own paths.
	Though interpretation of the phrase ``spreadsheet practice" can certainly vary
	from participant to participant, we readily accepted this interpretive
	flexibility since we wanted a range of views and backgrounds despite the
	persona. \par

@@

	\begin{wrapfigure}{L}{.25\textwidth} \centering \includegraphics{SUM}
		\label{fig:sum} \caption{An abundance of parameters in SUM} \end{wrapfigure}

@@

	\begin{wrapfigure}{R}{.25\textwidth} \centering \includegraphics[scale=.5]{longIF} \label{fig:longif}
		\caption{The many possibilities of IF, argument 3, demonstrate the problem of
			having too many options.} \end{wrapfigure}

@@

	[MOVED]
